---
title: 'My Paper'
author: 'You R. Name'
date: 2023-09-07
university: 'University of Toronto'
department: 'Your Department'
format:
  pdf:
    pdf-engine: pdflatex # xelatex and other also work
    highlight: tango # "default", "tango", "pygments", "kate", "monochrome", "espresso", "zenburn", "haddock"
    template: "../tex/template.tex"
    keep-tex: false
bibliography: ../bib/references.bib
csl: ../csl/chicago_manual_of_style_17th_edition_author_date.csl # Download your specific csl file and refer to it in the line below (see https://www.zotero.org/styles/)
fontsize: 12pt
linespacing: 1.5
margin: 1
paper: letterpaper # a4paper, executivepaper, legalpaper, etc also works
customfonts: false
sansserif: false
amsthm: true
outline: true
creativecommons: by-nc-sa # can be deleted
---

<!--
This text in between arrows is commented and will be ignored in the final PDF.

Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you receive a duplicate label error after knitting, make sure to delete the 
index.Rmd file and then knit again.

You can delete all the lines starting with "#" in the header, also you can 
delete the bibliography and/or csl as well
-->

<!--
This chunk sets the global options and can be deleted, you could add
warning = FALSE or any other R Markdown option
-->
```{r setup}
#| include: false

# You can delete the next lines, but render without removing it first.
# On Windows, you might need to restart your computer and try again if it doesn't work.
if (!require("knitr")) install.packages("knitr", repos = "https://cran.rstudio.com")
if (!require("ggplot2")) install.packages("ggplot2", repos = "https://cran.rstudio.com")
if (!require("stargazer")) install.packages("stargazer", repos = "https://cran.rstudio.com")

knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

<!--
The {#rmd-basics} text after the chapter declaration will allow us to link 
throughout the document back to the beginning of Chapter 1.  These labels will 
automatically be generated (if not specified) by changing the spaces to hyphens 
and capital letters to lowercase.  Look for the reference to this label at the 
beginning of Chapter 2.
-->

# R Markdown Basics {#rmd-basics}

Here is a brief introduction into using _R Markdown_. _Markdown_ is a simple 
formatting syntax for authoring HTML, PDF, and MS Word documents. _R Markdown_ 
provides the flexibility of _Markdown_ with the implementation of **R** input 
and output.  For more details on using _R Markdown_ see 
<https://rmarkdown.rstudio.com>.

Be careful with your spacing in _Markdown_ documents.  While whitespace largely 
is ignored, it does at times give _Markdown_ signals as to how to proceed.  As 
a habit, try to keep everything left aligned whenever possible, especially as 
you type a new paragraph.  In other words, there is no need to indent basic 
text in the Rmd document (in fact, it might cause your text to do funny things 
if you do).

## Lists

It's easy to create a list.  It can be unordered like

* Item 1
* Item 2

or it can be ordered like

1. Item 1
4. Item 2

Notice that I intentionally mislabeled Item 2 as number 4.  _Markdown_ 
automatically figures this out!  You can put any numbers in the list and it 
will create the list.  Check it out below.

To create a sublist, just indent the values a bit (at least four spaces or a 
tab).  (Here's one case where indentation is key!)

1. Item 1
1. Item 2
1. Item 3
    - Item 3a
    - Item 3b

## Line breaks

Make sure to add white space between lines if you'd like to start a new 
paragraph.  Look at what happens below in the outputted document if you don't:
Here is the first sentence.
Here is another sentence.
Here is the last sentence.

*Now for the correct way:* 

Here is the first sentence.

Here is another sentence.

Here is the last sentence.

## R chunks

When you click the **Knit** button above a document will be generated that 
includes both content as well as the output of any embedded **R** code chunks 
within the document. You can embed an **R** code chunk like this (`mtcars` is a 
built-in **R** dataset):

```{r mtcars}
summary(mtcars)
```

## Inline code

If you'd like to put the results of your analysis directly into your discussion,
add inline code like this:

> The `cos` of $2 \pi$ is `r cos(2*pi)`. 

Another example would be the direct calculation of the standard deviation:

> The standard deviation of `speed` in `cars` is `r sd(cars$speed)`.

One last neat feature is the use of the `ifelse` conditional statement which can
be used to output text depending on the result of an **R** calculation:

> `r ifelse(sd(cars$speed) < 6, "The standard deviation is less than 6.", "The 
> standard deviation is equal to or greater than 6.")`

Note the use of `>` here, which signifies a quotation environment that will be 
indented.

As you see with `$2 \pi$` above, mathematics can be added by surrounding the 
mathematical text with dollar signs.  More examples of this are in 
[Mathematical equations].

## Plots

Varsity blues already solves all the packages in order to insert plots right
away from your code.

```{r plots}
#| fig-width: 5
#| fig-height: 3
#| fig-cap: "An elementary plot"

library(ggplot2)

ggplot(mtcars) +
  geom_point(aes(x = cyl, y = wt, color = am))
```

## Tables

As for the case of plots, this package already solves all the dependencies in 
order to use different types of tables in \LaTeX.

### Simple table

```{r tables}
kable(xtabs(~ am, mtcars))
```

### Complex table (regression table)

```{r tables2}
#| results: 'asis'

library(stargazer)

model1 <- lm(mpg ~ cyl, mtcars)
model2 <- lm(mpg ~ cyl + am, mtcars)
model3 <- lm(mpg ~ cyl + am + wt, mtcars)

stargazer(model1, model2, model3, header = F, title = "Regression table")
```

\newpage

## Mathematical equations

Consider a function $f: U \to \mathbb{R}$, defined on an open set $U\subset\mathbb{R}$, is said to be **differentiable** at $a\in U$ if the derivative $f'(a) = \lim_{h \to 0}\frac{f(a+h)-f(a)}{h}$ exists. In general, $f$ is of class $\mathcal{C}^k$ if its first $k$ derivatives $f^{\prime}(x), f^{\prime\prime}(x), \ldots, f^{(k)}(x)$ exist and are continuous.

# Additional resources

- [_Markdown_ Cheatsheet](https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet)
- [_R Markdown_ Reference Guide](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)
- [_R Markdown_ Cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf)
- [_RStudio IDE_ Cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/rstudio-ide.pdf)
- [_RStudio IDE_ Official website](https://rstudio.com/products/rstudio/)
- [Introduction to `dplyr`](https://cran.rstudio.com/web/packages/dplyr/vignettes/dplyr.html)
- [`ggplot2` Documentation](https://ggplot2.tidyverse.org/)
- [`ggplot2` Cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf)

<!--
The bib chunk below must go last in this document according to how R Markdown 
renders.  More info is at 
http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
-->

\newpage

# References {-}

<!--
This manually sets the header for this unnumbered chapter.
-->
\markboth{References}{References}

<!--
To remove the indentation of the first entry.
-->
\noindent

<!--
This is just for testing with more citations for the bibliography at the end.  
Add other entries into the list here if you'd like them to appear in the 
bibliography even if they weren't explicitly cited in the document.
-->

---
nocite: | 
  @angel2000, @angel2001, @angel2002a
...
