---
title: 'My Assignment'
author: 'You R. Name'
university: 'University of Toronto'
department: 'Department of Statistical Sciences'
output:
  rmarkdown::pdf_document:
    latex_engine: pdflatex
    template: ../tex/template.tex
    # Supported styles include "default", "tango", "pygments", "kate", 
    # "monochrome", "espresso", "zenburn", "haddock". Delete the highlighting 
    # line NULL to turn off highlighting.
    highlight: tango
    keep_tex: yes
bibliography: ../bib/references.bib
# Download your specific csl file and refer to it in the line below (see https://www.zotero.org/styles/)
csl: ../csl/chicago.csl
fontsize: 12pt
linespacing: 1.5
---

<!--
This text in between arrows is commented and will be ignored in the final PDF.

Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you receive a duplicate label error after knitting, make sure to delete the 
index.Rmd file and then knit again.

You can delete all the lines starting with "#" in the header, also you can 
delete the bibliography and/or csl as well
-->

<!--
This chunk sets the global options and can be deleted, you could add
warning = FALSE or any other R Markdown option
-->
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

<!--
You can delete the next chunk afterwards, but Knit without removing it first to 
setup LaTeX. On Windows, you might need to restart your computer and try again 
if knitting the document doesn't work.
-->
```{r latex_setup}
if (!require("knitr")) 
  install.packages("knitr", repos = "https://cran.rstudio.com")
if (!require("rmarkdown")) 
  install.packages("rmarkdown", repos = "https://cran.rstudio.com")
if (!require("ggplot2")) 
  install.packages("ggplot2", repos = "https://cran.rstudio.com")
if (!require("stargazer")) 
  install.packages("stargazer", repos = "https://cran.rstudio.com")
if (!require("tinytex")) 
  install.packages("tinytex", repos = "https://cran.rstudio.com")

if (isFALSE(tinytex::is_tinytex())) tinytex::install_tinytex()
```

<!--
The {#rmd-basics} text after the chapter declaration will allow us to link 
throughout the document back to the beginning of Chapter 1.  These labels will 
automatically be generated (if not specified) by changing the spaces to hyphens 
and capital letters to lowercase.  Look for the reference to this label at the 
beginning of Chapter 2.
-->

# R Markdown Basics {#rmd-basics}

Here is a brief introduction into using _R Markdown_. _Markdown_ is a simple 
formatting syntax for authoring HTML, PDF, and MS Word documents. _R Markdown_ 
provides the flexibility of _Markdown_ with the implementation of **R** input 
and output.  For more details on using _R Markdown_ see 
<https://rmarkdown.rstudio.com>.  

Be careful with your spacing in _Markdown_ documents.  While whitespace largely 
is ignored, it does at times give _Markdown_ signals as to how to proceed.  As 
a habit, try to keep everything left aligned whenever possible, especially as 
you type a new paragraph.  In other words, there is no need to indent basic 
text in the Rmd document (in fact, it might cause your text to do funny things 
if you do).

## Lists

It's easy to create a list.  It can be unordered like

* Item 1
* Item 2

or it can be ordered like

1. Item 1
4. Item 2

Notice that I intentionally mislabeled Item 2 as number 4.  _Markdown_ 
automatically figures this out!  You can put any numbers in the list and it 
will create the list.  Check it out below.

To create a sublist, just indent the values a bit (at least four spaces or a 
tab).  (Here's one case where indentation is key!)

1. Item 1
1. Item 2
1. Item 3
    - Item 3a
    - Item 3b

## Line breaks

Make sure to add white space between lines if you'd like to start a new 
paragraph.  Look at what happens below in the outputted document if you don't:

Here is the first sentence.  Here is another sentence.  Here is the last 
sentence to end the paragraph.

This should be a new paragraph.

*Now for the correct way:* 

Here is the first sentence.  Here is another sentence.  Here is the last 
sentence to end the paragraph.

This should be a new paragraph.

## R chunks

When you click the **Knit** button above a document will be generated that 
includes both content as well as the output of any embedded **R** code chunks 
within the document. You can embed an **R** code chunk like this (`mtcars` is a 
built-in **R** dataset):

```{r mtcars}
summary(mtcars)
```

## Inline code

If you'd like to put the results of your analysis directly into your discussion,
add inline code like this:

> The `cos` of $2 \pi$ is `r cos(2*pi)`. 

Another example would be the direct calculation of the standard deviation:

> The standard deviation of `speed` in `cars` is `r sd(cars$speed)`.

One last neat feature is the use of the `ifelse` conditional statement which can
be used to output text depending on the result of an **R** calculation:

> `r ifelse(sd(cars$speed) < 6, "The standard deviation is less than 6.", "The 
> standard deviation is equal to or greater than 6.")`

Note the use of `>` here, which signifies a quotation environment that will be 
indented.

As you see with `$2 \pi$` above, mathematics can be added by surrounding the 
mathematical text with dollar signs.  More examples of this are in 
[Mathematical equations].

## Plots

Varsity blues already solves all the packages in order to insert plots right
away from your code.

```{r plots}
library(ggplot2)

ggplot(mtcars) +
  geom_point(aes(x = cyl, y = wt, color = am))
```

## Tables

As for the case of plots, this package alredy solves all the dependencies in 
order to use different types of tables in \LaTeX.

### Simple table

```{r tables}
kable(xtabs(~ am, mtcars))
```

### Complex table (regression table)

```{r tables2, results='asis'}
library(stargazer)

model1 <- lm(mpg ~ cyl, mtcars)
model2 <- lm(mpg ~ cyl + am, mtcars)
model3 <- lm(mpg ~ cyl + am + wt, mtcars)

stargazer(model1, model2, model3, header = F)
```

\newpage

# Mathematical equations

\TeX\ is the best way to typeset mathematics. Donald Knuth designed \TeX\ when 
he got frustrated at how long it was taking the typesetters to finish his book, 
which contained a lot of mathematics.

One nice feature of _R Markdown_ is its ability to read LaTeX code directly.

## A quick example of *some* of the package's shortcuts

Let $K$ be a field of \dfn{scalars}\index{field!scalar}\index{scalar}---usually 
either the real numbers $\R$ or the complex numbers 
$\mathbb C$, or occasionally the rationals $\mathbb Q$. A \idfn{vector space} 
over $K$ is a set
$V$ of \dfn{vectors}\index{vector} equipped with two operations, vector 
addition $(x,y)\mapsto x+y$, and scalar multiplication 
$(\alpha,x)\mapsto \alpha x$, where $x,y\in V$ and $\alpha\in K$.  The 
operations satisfy:

\begin{description}
\item[V.1] $x+y=y+x$
\item[V.2] $(x+y)+z=x+(y+z)$
\item[V.3] There is a vector $0$, satisfying $x+0=x$ for every vector $x$.
\item[V.4] $x+(-1)x=0$
\item[V.5] $\alpha(\beta x)=(\alpha \beta)x$
\item[V.6] $1x=x$
\item[V.7] $\alpha(x+y)=(\alpha x)+(\alpha y)$
\item[V.8] $(\alpha+\beta)x=(\alpha x)+(\beta x)$
\end{description}

## Statistical notation

* Algebra and semi-algebra: $\A,\S$
* Sigma field: $\F$
* Set of probability measures: $\prob$, $\prob{X}$, $\prob{A}$, etc
* Probability of $x$: $\P(X = x)$
* Different thetas: $\thetah, \thetat$
* Convergence in probability: $\thetah \probconv \theta$
* Union and intersection: $\infunion, \infintersection, \union{a}{b}, \intersection{a}{b}$
* Normal distributions: $\normalstd, \normal{\mu}{\sigma^2}$
* Measurable and probability space: $\mespace, \prspace$

## Calculus notation

Many shortcuts for derivatives. Let $\fn1$, then
$$\dfx, \dfy, \dfi, \dfj$$
$$\dFx, \dFy, \dFi, \dFj$$
Same for $g()$ which is $\gn1$
$$\dgx, \dgy, \dgi, \dgj$$

Now let $f$ be a $\C^2$ function
$$
\ddfyx, \ddfxy, \ddfx, \ddfy, \ddfij, \ddfji, \ldots, \dfij
$$

You can also have $\fgn1$ and $\fgnm$.

In general, you can write
$$
\prd{f}{x}
$$
or more general
$$\diff{L}{}{\beta}{} \text{ or } \diffat{L}{}{\beta}{}{\hat\beta_{\lambda}}$$
which is more flexible because it allows
$$\diffat{L}{2}{\beta}{3}{\hat\beta_{\lambda}}$$

## Matrices

You can also type matrices with relative efficiency. Here are some simple 
examples of use, and of course you can use \LaTeX default commands.

Start by replacing `\left[\begin{array}{...}` by `\vbmatrix{`. Note
that there is no column specifier. You can make as many columns as you like, but
they will all be centered. To finish, instead of `\end{array}\right]`, just type
`}`. Like this:

$$
\vbmatrix{\mbox{indices}&1&2&3&4\\
1&M_{1,1}&M_{1,2}&M_{1,3}&M_{1,4}\\
2&M_{2,1}&M_{2,2}&M_{2,3}&M_{2,4}
}
$$

You can also add a vertical rule and if you don’t like all the horizontal space around the vertical rule, you can get rid of it using plain TEX’s `\omit` command.

$$
\vbmatrix{\mbox{indices}&
1&2&\omit\vrule&3&4\\
1&M_{1,1}&M_{1,2}&\omit\vrule&
M_{1,3}&M_{1,4}\\
2&M_{2,1}&M_{2,2}&\omit\vrule&
M_{2,3}&M_{2,4}}
$$

If you think the vertical rule in the first row is too tall, you can shorten it 
using plain TEX’s `height` command.

$$
\vbmatrix{\mbox{indices}&
1&2&\omit\vrule height 1ex&3&4\\
1&M_{1,1}&M_{1,2}&\omit\vrule&
M_{1,3}&M_{1,4}\\
2&M_{2,1}&M_{2,2}&\omit\vrule&
M_{2,3}&M_{2,4}}
$$

Horizontal lines works as you might expect.

$$
\vbmatrix{\mbox{indices}&
1&2&\vrule&3&4\\
1&M_{1,1}&M_{1,2}&\vrule&
M_{1,3}&M_{1,4}\\\hline
2&M_{2,1}&M_{2,2}&\vrule&
M_{2,3}&M_{2,4}
}
$$

Notice how the `\hline` cuts across all the columns, but it doesn’t connect to 
the closing bracket. I am not sure I like this behavior, but you can use 
`\cline`.

$$
\vbmatrix{\mbox{indices}&1&2&
\vrule&3&4\\
1&M_{1,1}&M_{1,2}&\vrule&M_{1,3}&
M_{1,4}\\\cline{1-1} \cline{2-6}
2&M_{2,1}&M_{2,2}&\vrule&M_{2,3}&
M_{2,4}}
$$

$$
\vbmatrix{\mbox{indices}&1&2&
\vrule&3&4\\
1&M_{1,1}&M_{1,2}&\vrule&M_{1,3}&
M_{1,4}\\\cline{2-6}
2&M_{2,1}&M_{2,2}&\vrule&
M_{2,3}&M_{2,4}
}
$$

Here's a more complex example that summarises the previous steps.

$$
\vbmatrix{\text{indices}&
(1\cdot11)&(1\cdot12)&(1\cdot22)&
\vrule&
(2\cdot11)&(2\cdot12)&(2\cdot22)\\
%
(1)&\lambda(1)^2&2\lambda(1)\lambda(2)&\lambda(2)^2&\vrule&0&0&0\\
3
(2)&0&0&0&\vrule&\lambda(1)^2&2\lambda(1)\lambda(2)&\lambda(2)^2\\
\hline
(111)&3&0&0&\vrule&0&0&0\\
(112)&0&2&0&\vrule&1&0&0\\
(122)&0&0&1&\vrule&0&2&0\\
(222)&0&0&0&\vrule&0&0&3}
$$

where `\text{}` is defined in the amstext \LaTeX package.

### Changing delimiters

The varsitybluesmatrix package defines four style parameters that can be used to
change the appearance of the array. The first two are `\vbldelim` and 
`\vbrdelim`, the left and right delimiters. By default they are `[` and `]` but 
you can change them like this:

$$
\renewcommand{\vbldelim}{\{}
\renewcommand{\vbrdelim}{.}
\vbmatrix{\mbox{indices}&1&2&
\vrule&3&4\\
1&M_{1,1}&M_{1,2}&\vrule&M_{1,3}&
M_{1,4}\\\cline{2-6}
2&M_{2,1}&M_{2,2}&\vrule&
M_{2,3}&M_{2,4}
}
$$

Another option:

$$
\renewcommand{\vbldelim}{\langle}
\renewcommand{\vbrdelim}{|} 
\vbmatrix{\mbox{indices}&1&2&
\vrule&3&4\\
1&M_{1,1}&M_{1,2}&\vrule&M_{1,3}&
M_{1,4}\\\cline{2-6}
2&M_{2,1}&M_{2,2}&\vrule&
M_{2,3}&M_{2,4}
}
$$

Yet another option:

$$
\renewcommand{\vbldelim}{(}
\renewcommand{\vbrdelim}{)} 
\vbmatrix{\mbox{indices}&1&2&
\vrule&3&4\\
1&M_{1,1}&M_{1,2}&\vrule&M_{1,3}&
M_{1,4}\\\cline{2-6}
2&M_{2,1}&M_{2,2}&\vrule&
M_{2,3}&M_{2,4}
}
$$

### Changing the border row and column style

You can change the style of the border row and column column entries by 
redefining `\vbrowstyle` and `\vbcolstyle`, which are by default set to 
`\scriptstyle`. You could, for instance, say 
`\renewcommand{\vbrowstyle}{\relax}` to typeset the first row as usual. By the 
way, the upper left corner is governed by the column style. You can always use 
an `\mbox{}` to change the style of any particular entry.

### Changing spacing

Besides changing the delimiters, you can also change the space inserted after 
the first row and after the first column. These are governed by the lengths 
`\vbrowsep` and `\vbcolsep`. By default they are `0pt` and `.5\arraycolsep`, 
respectively.

$$
\setlength{\vbrowsep}{10pt}
\setlength{\vbcolsep}{10pt}
\vbmatrix{\mbox{indices}&1&2&
\vrule&3&4\\
1&M_{1,1}&M_{1,2}&\vrule&M_{1,3}&
M_{1,4}\\\cline{2-6}
2&M_{2,1}&M_{2,2}&\vrule&
M_{2,3}&M_{2,4}
}
$$

# Additional resources

- [_Markdown_ Cheatsheet](https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet)
- [_R Markdown_ Reference Guide](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)
- [_R Markdown_ Cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf)
- [_RStudio IDE_ Cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/rstudio-ide.pdf)
- [_RStudio IDE_ Official website](https://rstudio.com/products/rstudio/)
- [Introduction to `dplyr`](https://cran.rstudio.com/web/packages/dplyr/vignettes/dplyr.html)
- [`ggplot2` Documentation](https://ggplot2.tidyverse.org/)
- [`ggplot2` Cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf)

<!--
The bib chunk below must go last in this document according to how R Markdown 
renders.  More info is at 
http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
-->

# References {-}

<!--
This manually sets the header for this unnumbered chapter.
-->
\markboth{References}{References}

<!--
To remove the indentation of the first entry.
-->
\noindent

<!--
This is just for testing with more citations for the bibliography at the end.  
Add other entries into the list here if you'd like them to appear in the 
bibliography even if they weren't explicitly cited in the document.
-->

---
nocite: | 
  @angel2000, @angel2001, @angel2002a
...
