# R Markdown Basics {#rmd-basics}

Here is a brief introduction into using _R Markdown_. _Markdown_ is a simple 
formatting syntax for authoring HTML, PDF, and MS Word documents. _R Markdown_ 
provides the flexibility of _Markdown_ with the implementation of **R** input 
and output.  For more details on using _R Markdown_ see 
<https://rmarkdown.rstudio.com>.

Be careful with your spacing in _Markdown_ documents.  While whitespace largely 
is ignored, it does at times give _Markdown_ signals as to how to proceed.  As 
a habit, try to keep everything left aligned whenever possible, especially as 
you type a new paragraph.  In other words, there is no need to indent basic 
text in the Rmd document (in fact, it might cause your text to do funny things 
if you do).

## Lists

It's easy to create a list.  It can be unordered like

* Item 1
* Item 2

or it can be ordered like

1. Item 1
4. Item 2

Notice that I intentionally mislabeled Item 2 as number 4.  _Markdown_ 
automatically figures this out!  You can put any numbers in the list and it 
will create the list.  Check it out below.

To create a sublist, just indent the values a bit (at least four spaces or a 
tab).  (Here's one case where indentation is key!)

1. Item 1
1. Item 2
1. Item 3
    - Item 3a
    - Item 3b

## Line breaks

Make sure to add white space between lines if you'd like to start a new 
paragraph.  Look at what happens below in the outputted document if you don't:
Here is the first sentence.  Here is another sentence.  Here is the last 
sentence to end the paragraph.
This should be a new paragraph.

*Now for the correct way:* 

Here is the first sentence.  Here is another sentence.  Here is the last 
sentence to end the paragraph.

This should be a new paragraph.

## R chunks

When you click the **Knit** button above a document will be generated that 
includes both content as well as the output of any embedded **R** code chunks 
within the document. You can embed an **R** code chunk like this (`mtcars` is a 
built-in **R** dataset):

```{r mtcars}
summary(mtcars)
```

## Inline code

If you'd like to put the results of your analysis directly into your discussion,
add inline code like this:

> The `cos` of $2 \pi$ is `r cos(2*pi)`. 

Another example would be the direct calculation of the standard deviation:

> The standard deviation of `speed` in `cars` is `r sd(cars$speed)`.

One last neat feature is the use of the `ifelse` conditional statement which can
be used to output text depending on the result of an **R** calculation:

> `r ifelse(sd(cars$speed) < 6, "The standard deviation is less than 6.", "The 
> standard deviation is equal to or greater than 6.")`

Note the use of `>` here, which signifies a quotation environment that will be 
indented.

As you see with `$2 \pi$` above, mathematics can be added by surrounding the 
mathematical text with dollar signs.  More examples of this are in 
[Mathematical equations].

## Plots

Varsity blues already solves all the packages in order to insert plots right
away from your code.

```{r plots, fig.width=5, fig.height=3, fig.cap="An elementary plot"}
library(ggplot2)

ggplot(mtcars) +
  geom_point(aes(x = cyl, y = wt, color = am))
```

## Tables

As for the case of plots, this package already solves all the dependencies in 
order to use different types of tables in \LaTeX.

### Simple table

```{r tables}
kable(xtabs(~ am, mtcars))
```

### Complex table (regression table)

```{r tables2, results='asis'}
library(stargazer)

model1 <- lm(mpg ~ cyl, mtcars)
model2 <- lm(mpg ~ cyl + am, mtcars)
model3 <- lm(mpg ~ cyl + am + wt, mtcars)

stargazer(model1, model2, model3, header = F)
```

\newpage

## Mathematical equations

Consider a function $f: U \to \mathbb{R}$, defined on an open set $U\subset\mathbb{R}$, is said to be **differentiable** at $a\in U$ if the derivative $f'(a) = \lim_{h \to 0}\frac{f(a+h)-f(a)}{h}$ exists. In general, $f$ is of class $\mathcal{C}^k$ if its first $k$ derivatives $f^{\prime}(x), f^{\prime\prime}(x), \ldots, f^{(k)}(x)$ exist and are continuous.
